Mixed effect tweedie model in TMB

### 2020-06-24 analysis/tmb-sensor-explore/2-fit-temperature.Rmd
add good DO values from WCVI 2016 into models and predictions

Check predictions against raw data
```{r}
m_temp_rw <- readRDS(here::here("analysis/tmb-sensor-explore/models/model-temp-all-years-500kn-2.rds"))
predictions <- predict(m_temp_rw)
predictions$residuals <- residuals(m_temp_rw)
cor(predictions$est, predictions$temperature_c)^2 
```

@@

  g <- ggplot(predictions, aes_string("est", response)) +
    geom_point() +
    facet_wrap(~year) +
    coord_fixed() +
    geom_abline()
  print(g)

  g <- ggplot(predictions, aes(est, residuals)) +
    geom_point() +
    geom_smooth()
  print(g)

  g <- ggplot(predictions, aes_string(variable, "residuals")) +
    geom_point() +
    geom_smooth() +
    facet_wrap(~year)
  print(g)

  g <- ggplot(predictions, aes(X, Y, colour = residuals)) +
    geom_point() +
    coord_fixed() +
    scale_color_gradient2()+
    facet_wrap(~year)
  print(g)

@@

```{r }
m_temp_rw <- readRDS((here::here("analysis/tmb-sensor-explore/models/model-temp-all-years-500kn.rds")))
nd_all <- readRDS(here::here("analysis/VOCC/data/nd_all_synoptic.rds")) %>% filter(year<2019)

unique(nd_all$year)
glimpse(nd_all)
pred_temp <- predict(m_temp_rw, newdata = nd_all)

# saveRDS(pred_temp, file = "../VOCC/data/predicted_temp_allyears_revised2")
```

```{r}
p_nd <- readRDS("../VOCC/data/predicted_temp_allyears_revised2")

@@

(st_1n3 <- plot_map_raster(p_nd1, "epsilon_st") +
  ggtitle("Spatiotemporal random effects") +
  facet_wrap(~year) +
   scale_fill_gradient2( low = "Steel Blue 4",
 high = "Red 3", limits = c(-1.5,1.5), trans = sqrt,
    na.value = "red") +
  labs(fill = "epsilon") + 
  gfplot::theme_pbs() + 
  theme(
  legend.position = "none", 
  axis.title.x = element_blank(), axis.title.y = element_blank()))

@@

```{r}
p_nd4 <- p_nd %>% filter(ssid==4) #%>% filter(ssid!=4)
st_4 <- plot_map_raster(p_nd4, "epsilon_st") +
  #ggtitle(" ") +
  facet_wrap(~year) +
     scale_fill_gradient2( low = "Steel Blue 4",
 high = "Red 3", limits = c(-1.5,1.5), trans = sqrt,
    na.value = "red") +
          labs(fill = " ") + 
gfplot::theme_pbs() + theme(axis.title.x = element_blank(), axis.title.y = element_blank(), legend.position = c(.85,.15))
st_4
```

>>>>

Check predictions against raw data
```{r}
# m_temp_rw <- readRDS(here::here("analysis/tmb-sensor-explore/models/model-temp-all-years-500kn-2.rds"))
m_temp_rw <- readRDS(here::here("analysis/tmb-sensor-explore/models/model-temp-all-years-800kn.rds"))
predictions <- predict(m_temp_rw)
predictions$residuals <- residuals(m_temp_rw)
cor(predictions$est, predictions$temperature_c)^2 
sum(abs(predictions$residuals))

# 2908.879-2789.322
# 119.557/2908.879 # 4% reduction in abs errors
```

@@

  g <- ggplot(predictions, aes_string("est", response)) +
    geom_point(alpha = 0.2) +
    facet_wrap(~year) +
    coord_fixed() +
    geom_abline()
  print(g)

  g <- ggplot(predictions, aes(est, residuals)) +
    geom_point(alpha = 0.2) +
    geom_smooth()
  print(g)

  g <- ggplot(predictions, aes_string(variable, "residuals")) +
    geom_point(alpha = 0.2) +
    geom_smooth() +
    facet_wrap(~year)
  print(g)

  g <- ggplot(predictions, aes(X, Y, colour = residuals)) +
    geom_point(alpha = 0.2) +
    coord_fixed() +
    scale_color_gradient2()+
    facet_wrap(~year)
  print(g)

@@

```{r }
# m_temp_rw <- readRDS((here::here("analysis/tmb-sensor-explore/models/model-temp-all-years-500kn.rds")))
nd_all <- readRDS(here::here("analysis/VOCC/data/nd_all_synoptic.rds")) %>% filter(year<2019)

unique(nd_all$year)
glimpse(nd_all)
p_nd <- predict(m_temp_rw, newdata = nd_all)

saveRDS(p_nd, file = "../VOCC/data/predicted_temp_allyears_800kn.rds")
```

```{r}
# p_nd <- readRDS("../VOCC/data/predicted_temp_allyears_800kn.rds")

@@

(st_1n3 <- plot_map_raster(p_nd1, "epsilon_st") +
  ggtitle("Spatiotemporal random effects") +
  facet_wrap(~year) +
   scale_fill_gradient2( low = "Steel Blue 4",
 high = "Red 3", limits = c(-1.5,1.5), #trans = sqrt,
    na.value = "red") +
  labs(fill = "epsilon") + 
  gfplot::theme_pbs() + 
  theme(
  legend.position = "none", 
  axis.title.x = element_blank(), axis.title.y = element_blank()))

@@

```{r}
p_nd4 <- p_nd %>% filter(ssid==4) #%>% filter(ssid!=4)
st_4 <- plot_map_raster(p_nd4, "epsilon_st") +
  #ggtitle(" ") +
  facet_wrap(~year) +
     scale_fill_gradient2( low = "Steel Blue 4",
 high = "Red 3", limits = c(-1.5,1.5), #trans = sqrt,
    na.value = "red") +
          labs(fill = " ") + 
gfplot::theme_pbs() + theme(axis.title.x = element_blank(), axis.title.y = element_blank(), legend.position = c(.85,.15))
st_4
```

### 2020-06-24 analysis/tmb-sensor-explore/3-fit-do.Rmd
add good DO values from WCVI 2016 into models and predictions

  print("")
  print("AIC:")
  print(aic(m))
}
```

@@

@@

```{r}
ggplot(d_trawl, aes(temp, do_mlpl, colour = as.factor(ssid))) +
  geom_point() +
  facet_wrap(~year) +
  scale_color_viridis_d()
```

```{r}
ggplot(filter(d_trawl, year==2016), aes(temp, do_mlpl, colour = as.factor(ssid))) +
  geom_point() +
  #facet_wrap(~year) +
  scale_color_viridis_d(begin=0.65)
```

```{r}
ggplot(filter(d_trawl, year==2017), aes(temp, do_mlpl, colour = as.factor(ssid))) +
  geom_point() +
  #facet_wrap(~year) +
  scale_color_viridis_d(end = 0.35)
```

@@

Remove WCVI in 2016 due to extreme values and make 500 knot mesh
```{r}
d_trawl1 <- d_trawl %>% filter(!(year == 2016 & ssid ==4))

@@

saveRDS(do_model_all_DOY, file = (here::here(
  paste0("analysis/tmb-sensor-explore/models/model-do-without-wcvi2016-", 
    "800kn", #Sys.Date(), 
    ".rds")))
  )
```

```{r eval=FALSE}
do_model_all_DOY <- readRDS((here::here("analysis/tmb-sensor-explore/models/model-do-without-wcvi2016-800kn.rds")))
get_diag(do_model_all_DOY, response = "log_do", variable = "depth_scaled")
```

Predict DO for whole grid using quadratic of predicted temperature values and mean DOY
```{r eval=FALSE}
model1 <- readRDS(here::here("analysis/tmb-sensor-explore/models/model-temp-all-years-500kn.rds"))
model2 <- readRDS(here::here("analysis/tmb-sensor-explore/models/model-do-without-wcvi2016-800kn.rds"))

@@

# may want to filter estimates for depths less than 15 m
# pred_do <- pred_do %>% filter(depth>15)
saveRDS(pred_do, file=here::here(paste0("analysis/VOCC/data/predicted-DO-", Sys.Date(),".rds")))
```

Plot predictions from saved DO predictions file
```{r}
p_nd <- readRDS((here::here("analysis/VOCC/data/predicted-DO-2020-06-19.rds")))

>>>>

  print("")
  print("AIC:")
  print(aic(m))

}
```

@@

```{r}
ggplot(filter(d_trawl, depth_m_max < 80), aes(DOY, log_do, colour = depth)) +
  geom_point() +
  facet_wrap(~year) +
  scale_color_viridis_c()
ggplot(filter(d_trawl, depth_m_max > 80 & depth_m_max < 120), aes(DOY, log_do, colour = depth)) +
  geom_point() +
  facet_wrap(~year) +
  scale_color_viridis_c()
```

@@

```{r}
ggplot(filter(d_trawl, ssid == 4), aes(DOY,log_do,  colour = temp)) +
  geom_point() +
  facet_grid(ssid~year, scales = "free") +
  scale_color_viridis_c()
```

```{r}
ggplot(filter(d_trawl, ssid == 4), aes(DOY,temp,  colour = temp)) +
  geom_point() +
  facet_grid(ssid~year, scales = "free") +
  scale_color_viridis_c()
```

```{r}
ggplot(filter(d_trawl, ssid == 4 & !(year ==2016 & DOY > 149)), aes(X, Y, colour = log_do)) +
  geom_point() +
  facet_grid(ssid~year, scales = "free") +
  scale_color_viridis_c()
```

```{r}
ggplot(filter(d_trawl, year==2016), aes((temp), log(do_mlpl), colour = as.factor(ssid))) +
  geom_point() +
  #facet_wrap(~year) +
  scale_color_viridis_d(begin=0.65)
```

```{r}
ggplot(filter(d_trawl, year==2017), aes(x = (temp), log(do_mlpl), colour = as.factor(ssid))) +
  geom_point() +
  #facet_wrap(~year) +
  scale_color_viridis_d(end = 0.35)
```

@@

Remove most samples from WCVI in 2016 due to extreme values and make mesh
```{r}
d_trawl1 <- d_trawl %>% filter(!(year == 2016 & ssid ==4 & DOY > 149))

@@

saveRDS(do_model_all_DOY, file = (here::here(
  paste0("analysis/tmb-sensor-explore/models/model-do-without-wcvi2016b-", 
    "800kn", #Sys.Date(), 
    ".rds")))
  )
```

```{r eval=FALSE}
# do_model_all_DOY <- readRDS((here::here("analysis/tmb-sensor-explore/models/model-do-without-wcvi2016-depth3.rds")))
# do_model_all_DOY <- readRDS((here::here("analysis/tmb-sensor-explore/models/model-do-without-wcvi2016-depth2.rds")))
# do_model_all_DOY <- readRDS((here::here("analysis/tmb-sensor-explore/models/model-do-without-wcvi2016-800kn.rds")))
get_diag(do_model_all_DOY, response = "log_do", variable = "depth_scaled")
```

```{r}
predictions <- predict(do_model_all_DOY)
predictions$residuals <- residuals(do_model_all_DOY)
predictions$abs <- exp(abs(predictions$residuals))

good_rows <- filter (predictions, abs != "Inf")
sum(good_rows$abs)
19719.73-17459.49
2260.24/19719.73  # 11% reduction in abs errors
```

Predict DO for whole grid using quadratic of predicted temperature values and mean DOY
```{r eval=FALSE}
model1 <- readRDS(here::here("analysis/tmb-sensor-explore/models/model-temp-all-years-800kn.rds"))
model2 <- readRDS(here::here("analysis/tmb-sensor-explore/models/model-do-without-wcvi2016b-800kn.rds"))

@@

# may want to filter estimates for depths less than 15 m
# pred_do <- pred_do %>% filter(depth>15)
saveRDS(pred_do, file=here::here(paste0("analysis/VOCC/data/predicted-DO-", Sys.Date(),"-more2016.rds")))
```

Plot predictions from saved DO predictions file
```{r}
p_nd <- readRDS((here::here("analysis/VOCC/data/predicted-DO-2020-06-20-more2016.rds")))
